---
title: 'Practical Machine Learning: Prediction Exercise'
author: "Siddhesh Panhale"
date: "9/6/2020"
output: html_document
---
#Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

#Data Loading and Exploratory Analysis

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

```{r}
library(dplyr)
library(ggplot2)
library(knitr)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(RColorBrewer)
library(RGtk2)
library(gbm)

training_Url<- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testing_Url<- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training_data<- read.csv(url(training_Url))
testing_data<- read.csv(url(testing_Url))
dim(training_data)
```

```{r}
dim(testing_data)
```

#Both datasets has 160 variables that has some NAs, that can be cleaned as below. This removes Near Zero variance(NZV) variables too.
```{r}
nzv <- nearZeroVar(training_data)

train_data <- training_data[,-nzv]
test_data <- testing_data[,-nzv]
dim(train_data)
```

```{r}
dim(test_data)
```
```{r}
na_val_col <- sapply(train_data, function(x) mean(is.na(x))) > 0.95
train_data <- train_data[,na_val_col == FALSE]
test_data <- test_data[,na_val_col == FALSE]
dim(train_data)
```

```{r}
dim(test_data)  
```

#Removing non numeric variables
```{r}
train_data<- train_data[, 8:59]
test_data<- test_data[, 8:59]
dim(train_data)
```

```{r}
dim(test_data)
```

##Data Partioning
#Partitioning the data into two parts: 60% Training and 40% Testing
```{r}
inTrain<- createDataPartition(train_data$classe, p=0.6, list=FALSE)
inTrain<- createDataPartition(train_data$classe, p=0.6, list=FALSE)
training<- train_data[inTrain,]
testing<- train_data[-inTrain,]
dim(training)
```

```{r}
dim(testing)
```

##Prediction Model Building

#Decision Tree
```{r}
DT_model<- train(classe ~. , data=training, method= "rpart")
fancyRpartPlot(DT_model$finalModel)
```
```{r}
set.seed(21243)
DT_prediction<- predict(DT_model, testing)
confusionMatrix(table(DT_prediction, testing$classe)) 
```
#Decision Tree Model we see the prediction accuracy is 48% which is not upto satisfactory level

#Random Forest MOdel
```{r}
set.seed(26817)
RF_model<- train(classe ~. , data=training, method= "rf", ntree=10)
RF_prediction<- predict(RF_model, testing)
RF_cm<-confusionMatrix(table(RF_prediction, testing$classe))
RF_cm
```

```{r}
plot(RF_cm$table, col=RF_cm$byClass, main="Random Forest Accuracy")
```
#From above we see that Random Forest Model provides accuracy of 99% which is highly satisified.

<!-- ##Gradient Boosting Model -->
<!-- ```{r} -->
<!-- set.seed(9999) -->
<!-- gbm_model<- train(classe~., data=training, method="gbm", verbose= FALSE) -->
<!-- gbm_model$finalmodel -->
<!-- ``` -->

<!-- ```{r} -->
<!-- gbm_prediction<- predict(gbm_model, testing) -->
<!-- gbm_cm<-confusionMatrix(gbm_prediction, testing$classe) -->
<!-- gbm_cm -->
<!-- ``` -->
<!-- #Gradient Boosting Model we see the prediction accuracy is 96% which is satisfied -->
<!-- ```{r} -->
<!-- RF_cm$overall -->
<!-- ``` -->

<!-- ```{r} -->
<!-- gbm_cm$overall -->
<!-- ``` -->

#Gradient Boosting Model
```{r}
set.seed(12345)
controlGBM <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
GBM_model  <- train(classe ~ ., data=training, method = "gbm",
                    trControl = controlGBM, verbose = FALSE)
GBM_model$finalModel
```

```{r}
GBM_prediction <- predict(GBM_model, newdata=testing)
GBM_cm <- confusionMatrix(table(GBM_prediction, testing$classe))
GBM_cm
```
##From Gradient Boosting Model, we see that the prediction accuracy is 96% which is satisfied.

```{r}
plot(GBM_cm$table, col = GBM_cm$byClass, 
     main = paste("GBM - Accuracy =", round(GBM_cm$overall['Accuracy'], 4)))
```

##compairing both models
```{r}
RF_cm$overall
```

```{r}
GBM_cm$overall
```

##CONCLUSION
#We see that, Random Forest is more accurate than Gradient Boosting Model at upto 99% of accuracy level.

##Prediction using Random Forest Model on testing data.
```{r}
prediction_test<- predict(RF_model, test_data)
prediction_test
```